---
title: "Part 1: Introduction to Correlation & Regression"
subtitle: "Lab 4: Linear Models"
format: html
filters:
  - webr
webr:
  packages: []
  show-startup-message: false
---

```{r, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Watch the Demo

Follow along with the slides below. You can advance slides with the arrow keys or click the arrows.

```{=html}
<iframe src="part1-slides.html" width="100%" style="aspect-ratio: 16/9; border: 1px solid #ccc; border-radius: 8px;"></iframe>
```

::: {.callout-tip}
## Tip
Click inside the slide area first, then use arrow keys to navigate.
:::

---

## Try It Yourself

Now practice what you just learned! The code cells below are **interactive**.

::: {.callout-tip}
## How to Practice
1. **Click "Run Code"** to execute the code and see the result
2. **Edit the code directly** - change values and see what happens
3. **Experiment freely** - you can't break anything!
:::

### Types of Relationships

Visualise positive, negative, and no relationship:

```{webr-r}
# Positive relationship
set.seed(42)
x <- 1:20
y_pos <- x + rnorm(20, 0, 3)
plot(x, y_pos, main = "Positive Relationship",
     xlab = "X", ylab = "Y", pch = 19, col = "steelblue")
abline(lm(y_pos ~ x), col = "red", lwd = 2)
```

::: {.callout-note}
## Challenge
Change `y_pos <- x + rnorm(20, 0, 3)` to `y_neg <- -x + rnorm(20, 0, 3)` and update the plot to show a negative relationship.
:::

### Calculating r by Hand (Conceptually)

```{webr-r}
x <- c(2, 4, 6, 8, 10)
y <- c(3, 5, 4, 8, 9)

r <- cor(x, y)
cat("Correlation r =", round(r, 3), "\n\n")

cat("Interpreting r =", round(r, 3), ":\n")
size <- if (abs(r) < 0.1) "Negligible" else if (abs(r) < 0.3) "Small" else if (abs(r) < 0.5) "Medium" else "Large"
cat(size, "relationship\n")

direction <- if (r > 0) "Positive" else "Negative"
cat("Direction:", direction)
```

### r² — Shared Variance

```{webr-r}
r <- 0.70
r_squared <- r^2

cat("r =", r, "\n")
cat("r² =", round(r_squared, 3), "\n\n")
cat("Interpretation: The two variables share",
    round(r_squared * 100, 1), "% of their variance.\n\n")
cat("The remaining", round((1 - r_squared) * 100, 1),
    "% of variance in Y is\nexplained by other factors.")
```

---

## Self-Assessment

Check your understanding before moving on.

::: {.webex-check .webex-box}

**1. Correlation and regression are used when both variables are:**

`r mcq(c("Categorical", answer = "Continuous", "One categorical, one continuous", "Ordinal"))`

**2. Pearson's r ranges from:**

`r mcq(c("0 to 1", "0 to 100", answer = "-1 to +1", "-∞ to +∞"))`

**3. r = -0.80 indicates a:**

`r mcq(c("Weak positive relationship", "No relationship", "Weak negative relationship", answer = "Strong negative relationship"))`

**4. r² tells us the proportion of `r fitb("variance", width = 10)` shared between two variables.**

**5. True or False: Correlation implies causation.**

`r torf(FALSE)`

:::

---

## Download Exercise

Ready to practice in RStudio? Follow these steps:

### ![](/assets/images/svg/download.svg){width=24} Step 1: Download the File

```{r}
#| echo: false
downloadthis::download_file(
  path = "../../templates/lab4-exercise.qmd",
  output_name = "lab4-exercise",
  button_label = "Download Exercise (.qmd)",
  button_type = "warning",
  icon = "fa fa-download",
  class = "btn-lg"
)
```

### Step 2: Open RStudio Server

1. Go to [RStudio Server](https://psy909.gold.ac.uk/rstudioserver/)
2. Log in with `yourusername@campus.goldsmiths.ac.uk`

### Step 3: Create Folder and Upload

1. Create a folder called `RLab4`
2. Upload `lab4-exercise.qmd` to the folder
3. Open the file and change your name
4. Click Render to test it works!

---

## Next Steps

::: {.d-flex .justify-content-between}
[← Back to Lab 4 Overview](../../labs/lab04.qmd){.btn .btn-secondary}

[Continue to Part 2: Correlation →](part2.qmd){.btn .btn-primary}
:::
