---
title: "Part 1: Introduction to T-Tests"
subtitle: "Lab 2: T-Tests"
format: html
filters:
  - webr
webr:
  packages: []
  show-startup-message: false
---

```{r, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Watch the Demo

Follow along with the slides below. You can advance slides with the arrow keys or click the arrows.

```{=html}
<iframe src="part1-slides.html" width="100%" height="500px" style="border: 1px solid #ccc; border-radius: 8px;"></iframe>
```

::: {.callout-tip}
## Tip
Click inside the slide area first, then use arrow keys to navigate.
:::

---

## Try It Yourself

Now practice what you just learned! The code cells below are **interactive** - you can edit them directly on this page.

::: {.callout-tip}
## How to Practice
1. **Click "Run Code"** to execute the code and see the result
2. **Edit the code directly** - click inside the code box, change numbers or add new lines, then click "Run Code" again
3. **Experiment freely** - try different values and see what happens!

You're running R code right here in your browser - no need to switch to RStudio yet!
:::

### Understanding Sample Means

Every sample has a mean, but how do we know if it's meaningfully different from what we expect?

```{webr-r}
# Create a sample of 20 test anxiety scores
set.seed(42)
anxiety_scores <- round(rnorm(20, mean = 55, sd = 10))

# Calculate the sample mean
mean(anxiety_scores)
```

::: {.callout-note}
## Challenge
**Edit the code above** to change the sample size to 50. Does the mean change much?
:::

### Comparing to a Population Mean

If the population mean is 50, is our sample mean of ~55 really different?

```{webr-r}
# Our sample
set.seed(42)
anxiety_scores <- round(rnorm(20, mean = 55, sd = 10))

# Compare to population mean of 50
sample_mean <- mean(anxiety_scores)
pop_mean <- 50

# Difference
sample_mean - pop_mean
```

### The Key Question

Is this difference:
- **Real** (a true difference between our sample and the population)?
- **Just chance** (random sampling variation)?

**This is what a t-test helps us determine!**

```{webr-r}
# Preview of what we'll learn
set.seed(42)
anxiety_scores <- round(rnorm(20, mean = 55, sd = 10))

# A one-sample t-test comparing to mu = 50
t.test(anxiety_scores, mu = 50)
```

::: {.callout-note}
## What to Notice
Look for:
- **t value**: How many standard errors away from the population mean
- **p-value**: Probability of seeing this result if null hypothesis is true
- **confidence interval**: Range of plausible population means
:::

---

## Self-Assessment

Check your understanding before moving on.

::: {.webex-check .webex-box}

**1. What does a t-test help us determine?**

`r mcq(c("The exact population mean", answer = "Whether a sample difference is likely due to chance", "The sample size needed", "The standard deviation"))`

**2. The null hypothesis typically states that there is:**

`r mcq(c(answer = "No difference", "A large difference", "A positive difference", "A negative difference"))`

**3. A p-value less than 0.05 means the result is considered statistically `r fitb("significant")`.**

**4. True or False: A t-test can only compare two groups.**

`r torf(FALSE)`

**5. Which t-test compares a sample mean to a known population value?**

`r mcq(c(answer = "One-sample t-test", "Independent samples t-test", "Paired samples t-test", "Chi-square test"))`

:::

---

## Download Exercise

Ready to practice in RStudio? Follow these steps carefully:

### ![](/assets/images/svg/download.svg){width=24} Step 1: Download the File

Click the button below to download the exercise notebook:

```{r}
#| echo: false
downloadthis::download_file(
  path = "../../templates/lab2-exercise.qmd",
  output_name = "lab2-exercise",
  button_label = "Download Exercise (.qmd)",
  button_type = "warning",
  icon = "fa fa-download",
  class = "btn-lg"
)
```

::: {.callout-important}
## Where Did It Go?

The file `lab2-exercise.qmd` will download to your **Downloads folder** (or wherever your browser saves downloads).

**Can't find it?**
- Check your browser's download bar (usually at the bottom)
- Look in your Downloads folder: `~/Downloads/` (Mac) or `C:\Users\YourName\Downloads\` (Windows)
:::

### Step 2: Open RStudio Server

1. Go to [RStudio Server](https://psy909.gold.ac.uk/rstudioserver/)
2. Log in with your Goldsmiths credentials

::: {.callout-important}
## Login Format

**Important:** Use your full campus email address format:

- Format: `yourusername@campus.goldsmiths.ac.uk`
- Example: `jbond007@campus.goldsmiths.ac.uk`
:::

### Step 3: Create a Project Folder

1. In RStudio, look at the **Files** tab in the bottom-right pane
2. Click **New Folder** (or right-click and select "New Folder")
3. Name it `RLab2`
4. **Double-click** the folder to open it

### Step 4: Upload the File

1. Make sure you're **inside** your `RLab2` folder
2. Click the **Upload** button (looks like an arrow pointing up)
3. Click **Choose File** and navigate to your Downloads folder
4. Select `lab2-exercise.qmd`
5. Click **OK** to upload

### Step 5: Open the File

1. In the Files pane, **click** on `lab2-exercise.qmd`
2. The file will open in the **Source** pane (top-left)
3. You're ready to start working!

---

## Your First Task: Change Your Name and Render

Now that you have the notebook open:

1. **Find line 4** where it says `author: "YOUR NAME HERE"`
2. **Replace** with your actual name (keep the quotation marks!)
3. **Click the Render button** (top toolbar)
4. Check that your name appears in the output!

::: {.callout-success}
## Success!

If you see your name in the rendered HTML document, you're ready for Part 2!
:::

---

## Next Steps

::: {.d-flex .justify-content-between}
[← Back to Lab 2 Overview](../../labs/lab02.qmd){.btn .btn-secondary}

[Continue to Part 2: One-Sample T-Test →](part2.qmd){.btn .btn-primary}
:::
