---
title: "Week 12 Data Analysis"
subtitle: "Data Skills Assignment"
author: "Your Name"
date: "`r Sys.Date()`"
format: 
  html:
    toc: false
    number-sections: true
    highlight-style: github
    code-fold: true
    code-summary: "Show code"
execute:
  echo: false
  warning: false
  message: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 8, fig.height = 5)
```

## Introduction

This document presents the analysis of Week 12 reaction time data. The dataset contains measurements from 60 participants across baseline, control, and experimental conditions, with participants assigned to either Happy or Scary experimental conditions.

## Loading the Data

```{r load-data}
# Load the data directly from GitHub CSV
data_url <- "https://raw.githubusercontent.com/LittleMonkeyLab/datarepo/main/Y1_Week_12_data.csv"

# Load the data using base R
Lab_12_data <- read.csv(data_url, stringsAsFactors = FALSE)

# Convert to tibble for better display
library(tibble)
Lab_12_data <- as_tibble(Lab_12_data)

# Display data structure
cat("Data loaded successfully!\n")
cat("Dataset dimensions:", nrow(Lab_12_data), "participants ×", ncol(Lab_12_data), "variables\n\n")
head(Lab_12_data)
```

The dataset contains reaction time measurements with the following structure:

- **ID**: Participant identifier
- **Condition**: Experimental condition (1 = Happy, 2 = Scary)
- **Base1, Base2, Base3**: Three baseline condition trials
- **Control1, Control2, Control3**: Three control condition trials  
- **Expt1, Expt2, Expt3**: Three experimental condition trials

## Task 1: Calculate Overall Means

Calculate the mean for baseline, control, and experimental conditions (averaging across all trials and participants).

```{r task1-calculate}
# Calculate column means first
numeric_cols <- sapply(Lab_12_data, is.numeric)
myColumnMeans <- colMeans(Lab_12_data[, numeric_cols, drop = FALSE], na.rm = TRUE)

# Find columns for each condition
base_cols <- grep("^Base", names(Lab_12_data), ignore.case = TRUE)
control_cols <- grep("^Control", names(Lab_12_data), ignore.case = TRUE)
expt_cols <- grep("^Expt", names(Lab_12_data), ignore.case = TRUE)

# Calculate overall means for each condition (averaging across trials)
baseMean <- mean(myColumnMeans[base_cols], na.rm = TRUE)
controlMean <- mean(myColumnMeans[control_cols], na.rm = TRUE)
expMean <- mean(myColumnMeans[expt_cols], na.rm = TRUE)

cat("Overall Means:\n")
cat("==============\n\n")
cat("Baseline mean:", round(baseMean, 2), "\n")
cat("Control mean:", round(controlMean, 2), "\n")
cat("Experimental mean:", round(expMean, 2), "\n")

# Store for later use
overall_means <- c(baseMean, controlMean, expMean)
```

**What pattern do you notice in these means?**

Write your observations here:

---

## Task 2: Calculate Means by Condition

Calculate the mean for baseline, control, and experimental, **separated by condition** (1=happy, 2=scary).

```{r task2-calculate}
# Use by() function to calculate means by condition
groupedMeans <- by(Lab_12_data, Lab_12_data$Condition, colMeans, na.rm = TRUE)

# Extract means for each group
gp1Means <- groupedMeans[[1]]  # Condition 1 (Happy)
gp2Means <- groupedMeans[[2]]  # Condition 2 (Scary)

# Calculate means for each condition type within each group
gp1BaseMean <- mean(gp1Means[base_cols], na.rm = TRUE)
gp1ControlMean <- mean(gp1Means[control_cols], na.rm = TRUE)
gp1ExpMean <- mean(gp1Means[expt_cols], na.rm = TRUE)

gp2BaseMean <- mean(gp2Means[base_cols], na.rm = TRUE)
gp2ControlMean <- mean(gp2Means[control_cols], na.rm = TRUE)
gp2ExpMean <- mean(gp2Means[expt_cols], na.rm = TRUE)

cat("Means by Experimental Condition:\n")
cat("================================\n\n")
cat("Happy Condition (1):\n")
cat("  Baseline:", round(gp1BaseMean, 2), "\n")
cat("  Control:", round(gp1ControlMean, 2), "\n")
cat("  Experimental:", round(gp1ExpMean, 2), "\n\n")

cat("Scary Condition (2):\n")
cat("  Baseline:", round(gp2BaseMean, 2), "\n")
cat("  Control:", round(gp2ControlMean, 2), "\n")
cat("  Experimental:", round(gp2ExpMean, 2), "\n")

# Store for plotting
happy_means <- c(gp1BaseMean, gp1ControlMean, gp1ExpMean)
scary_means <- c(gp2BaseMean, gp2ControlMean, gp2ExpMean)
```

**What differences do you notice between the happy and scary conditions?**

Write your observations here:

---

## Task 3: Create a Single Line Graph

Plot a line graph showing the overall means across the three conditions (Baseline → Control → Experimental).

```{r task3-plot}
# Use the calculated means from Task 1
conditions <- c("Baseline", "Control", "Experimental")

plot(1:3, overall_means, 
     type = "b",  # Both points and lines
     xlab = "Condition", 
     ylab = "Mean Reaction Time (ms)",
     main = "Overall Mean Reaction Times Across Conditions",
     xaxt = "n",
     pch = 19,
     col = "#9B1B30",
     lwd = 2,
     ylim = c(min(overall_means) * 0.95, max(overall_means) * 1.05))
axis(1, at = 1:3, labels = conditions)
grid()
```

**What does this graph suggest about the pattern of reaction times?**

Write your interpretation here:

---

## Task 4: Create a Multi-Line Graph

Plot a line graph with **2 lines** (one for each experimental condition: happy vs. scary) on the same graph.

```{r task4-plot}
# Use the calculated means from Task 2
all_means <- c(happy_means, scary_means)
y_range <- c(min(all_means) * 0.95, max(all_means) * 1.05)

plot(1:3, happy_means, 
     type = "b",
     xlab = "Condition", 
     ylab = "Mean Reaction Time (ms)",
     main = "Mean Reaction Times by Experimental Condition",
     xaxt = "n",
     pch = 19,
     col = "#2E7D32",  # Green for happy
     lwd = 2,
     ylim = y_range)
lines(1:3, scary_means, 
      type = "b",
      pch = 19,
      col = "#C62828",  # Red for scary
      lwd = 2)
axis(1, at = 1:3, labels = conditions)
legend("topright", 
       legend = c("Happy", "Scary"),
       col = c("#2E7D32", "#C62828"),
       lty = 1,
       lwd = 2,
       pch = 19)
grid()
```

**What does this comparison reveal about the differences between conditions?**

#Question_1.1 - REPLACE THIS TEXT WITH YOUR FINDINGS

---

## Summary


#Question_1.2 - REPLACE THIS TEXT WITH YOUR SUMMARY OF THE MAIN FINDINGS.
