---
title: "Lab 1 Follow-On: Transfer Practice"
subtitle: "Applying Your Skills to New Data"
author: "YOUR NAME HERE"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    code-fold: false
    code-annotations: below
    theme: cosmo
    embed-resources: true
---

```{r}
#| label: setup
#| message: false
#| warning: false

# Load required packages
library(readr)
library(dplyr)
library(ggplot2)
```

# Introduction

This follow-on exercise uses a **different dataset** to help you consolidate your Lab 1 skills. You'll apply the same techniques to new data with less scaffolding.

::: {.callout-note}
## Dataset: Wellbeing Study

Researchers investigated the effect of a mindfulness intervention on psychological wellbeing. Participants completed a 6-item wellbeing questionnaire before and after an 8-week programme.

- **Participants**: 120 university students
- **Design**: 2 (group: control vs mindfulness) x 2 (time: pre vs post)
- **Measures**: 6 wellbeing items (w1-w6) on a 1-7 scale
- **Reverse-scored items**: w3 and w5
:::


# Task 1: Load and Inspect

Load the wellbeing dataset and explore its structure.

```{r}
#| label: load-data

# Load data from GitHub
data_url <- "https://raw.githubusercontent.com/LittleMonkeyLab/datarepo/main/wellbeing_data.csv"
wellbeing <- read_csv(data_url, show_col_types = FALSE)
```

```{r}
#| label: inspect

# YOUR CODE: View the first few rows


# YOUR CODE: Check the structure


# YOUR CODE: Check dimensions

```

::: {.callout-note}
## Questions

1. How many participants? ___
2. How many variables? ___
3. What are the condition levels? ___
:::

# Task 2: Reverse Scoring

Items w3 and w5 are negatively worded. Reverse-score them.

::: {.callout-tip}
## Hint

For a 1-7 scale, the formula is: `reversed = 8 - original`
:::

```{r}
#| label: reverse-score

# YOUR CODE: Create reversed versions of w3 and w5


# YOUR CODE: Verify the reversal worked

```

# Task 3: Create Total Score

Sum all items (using reversed versions for w3 and w5) to create `wellbeing_total`.

```{r}
#| label: total-score

# YOUR CODE: Create the total wellbeing score


# YOUR CODE: View a few rows to check

```

# Task 4: Descriptive Statistics

## Overall Statistics

```{r}
#| label: overall-stats

# YOUR CODE: Calculate overall mean, SD, and n


```

## Grouped Statistics

```{r}
#| label: grouped-stats

# YOUR CODE: Calculate mean, SD, and n by condition and timepoint


```

::: {.callout-note}
## Questions

Fill in from your output:

| Condition | Timepoint | M | SD | n |
|-----------|-----------|---|----|----|
| control | pre | ___ | ___ | ___ |
| control | post | ___ | ___ | ___ |
| mindfulness | pre | ___ | ___ | ___ |
| mindfulness | post | ___ | ___ | ___ |
:::

# Task 5: Visualisation

Create a boxplot showing wellbeing scores by condition and timepoint.

```{r}
#| label: boxplot
#| fig-width: 8
#| fig-height: 5

# YOUR CODE: Create a boxplot


```

Create one additional visualisation of your choice:

```{r}
#| label: custom-viz
#| fig-width: 8
#| fig-height: 5

# YOUR CODE: Create another visualisation


```

# Task 6: APA Write-Up

Write an APA-style paragraph describing your findings. Include:

- Overall mean and SD
- Means and SDs by condition at each timepoint
- Description of any apparent effects

**YOUR WRITE-UP:**

[Write your paragraph here]

# Task 7: Comparison

::: {.callout-note}
## Reflection

Compare this analysis to Lab 1:

1. **What was the same?**

[YOUR ANSWER]

2. **What was different?**

[YOUR ANSWER]

3. **What formula did you use for reverse-scoring, and why was it different from Lab 1?**

[YOUR ANSWER]

4. **How confident do you feel applying these skills to new data?**

[YOUR ANSWER]
:::

# Solutions

::: {.callout-caution collapse="true"}
## Click to reveal solutions (try first!)

### Task 2: Reverse Scoring
```r
wellbeing$rev_w3 <- 8 - wellbeing$w3
wellbeing$rev_w5 <- 8 - wellbeing$w5
```

### Task 3: Total Score
```r
wellbeing$wellbeing_total <- wellbeing$w1 + wellbeing$w2 +
                              wellbeing$rev_w3 + wellbeing$w4 +
                              wellbeing$rev_w5 + wellbeing$w6
```

### Task 4: Grouped Statistics
```r
wellbeing |>
  group_by(condition, timepoint) |>
  summarise(
    mean = mean(wellbeing_total, na.rm = TRUE),
    sd = sd(wellbeing_total, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )
```

### Task 5: Boxplot
```r
ggplot(wellbeing, aes(x = condition, y = wellbeing_total, fill = timepoint)) +
  geom_boxplot(alpha = 0.7) +
  theme_minimal() +
  labs(
    title = "Wellbeing Scores by Condition and Timepoint",
    x = "Condition",
    y = "Wellbeing Score",
    fill = "Timepoint"
  )
```
:::

# Session Info

```{r}
#| label: session-info

sessionInfo()
```
