---
title: "Part 5: APA Write-Up"
subtitle: "Lab 1: Descriptive Statistics"
format: html
filters:
  - webr
webr:
  packages: ['readr', 'dplyr']
  show-startup-message: false
---

```{r, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Watch the Demo

Follow along with the slides below.

```{=html}
<iframe src="part5-slides.html" width="100%" height="500px" style="border: 1px solid #ccc; border-radius: 8px;"></iframe>
```

::: {.callout-tip}
## Tip
Click inside the slide area first, then use arrow keys to navigate.
:::

---

## Try It Yourself

::: {.callout-tip}
## How to Practice
1. **Generate your summary statistics**
2. **Use the template** to write your paragraph
3. **Copy to your RStudio exercise** for submission
:::

### Generate Your Statistics

Run this to get all the numbers you need:

```{webr-r}
library(readr)
library(dplyr)

# Load and prepare data
data_url <- "https://raw.githubusercontent.com/LittleMonkeyLab/datarepo/main/Y1_Week13_Lab1_grit_data.csv"
data <- read_csv(data_url, show_col_types = FALSE)

# Reverse-score items
data$rev_G1 <- 6 - data$G1
data$rev_G3 <- 6 - data$G3
data$rev_G5 <- 6 - data$G5
data$rev_G6 <- 6 - data$G6

# Create subscales
data$consistency <- data$rev_G1 + data$rev_G3 + data$rev_G5 + data$rev_G6
data$perseverance <- data$G2 + data$G4 + data$G7 + data$G8
data$Study_Year <- factor(data$Study_Year)

# Overall statistics
cat("=== OVERALL ===\n")
cat("Consistency of Interest:\n")
cat("M =", round(mean(data$consistency), 2), "\n")
cat("SD =", round(sd(data$consistency), 2), "\n\n")
cat("Perseverance of Effort:\n")
cat("M =", round(mean(data$perseverance), 2), "\n")
cat("SD =", round(sd(data$perseverance), 2), "\n")
cat("N =", length(data$consistency), "\n\n")

# By Study Year
cat("=== BY STUDY YEAR ===\n")
summary_stats <- data |>
  group_by(Study_Year) |>
  summarise(
    M_consistency = round(mean(consistency), 2),
    SD_consistency = round(sd(consistency), 2),
    M_perseverance = round(mean(perseverance), 2),
    SD_perseverance = round(sd(perseverance), 2),
    n = n(),
    .groups = "drop"
  )

print(summary_stats)
```

### Create a Formatted Table

```{webr-r}
library(readr)
library(dplyr)

data_url <- "https://raw.githubusercontent.com/LittleMonkeyLab/datarepo/main/Y1_Week13_Lab1_grit_data.csv"
data <- read_csv(data_url, show_col_types = FALSE)
data$rev_G1 <- 6 - data$G1
data$rev_G3 <- 6 - data$G3
data$rev_G5 <- 6 - data$G5
data$rev_G6 <- 6 - data$G6
data$consistency <- data$rev_G1 + data$rev_G3 + data$rev_G5 + data$rev_G6
data$perseverance <- data$G2 + data$G4 + data$G7 + data$G8
data$Study_Year <- factor(data$Study_Year)

summary_stats <- data |>
  group_by(Study_Year) |>
  summarise(
    M_consistency = round(mean(consistency), 2),
    SD_consistency = round(sd(consistency), 2),
    M_perseverance = round(mean(perseverance), 2),
    SD_perseverance = round(sd(perseverance), 2),
    n = n(),
    .groups = "drop"
  )

knitr::kable(summary_stats,
             caption = "Table 1. Descriptive Statistics for Grit Subscales by Study Year",
             col.names = c("Study Year", "M (Consistency)", "SD (Consistency)", 
                          "M (Perseverance)", "SD (Perseverance)", "n"))
```

---

## APA Write-Up Template

Use this template for your results paragraph. Fill in the blanks with your statistics:

::: {.callout-note}
## Template

Participants reported [describe level] levels of Consistency of Interest overall (*M* = ___, *SD* = ___) and Perseverance of Effort (*M* = ___, *SD* = ___). First-year students showed [description] Consistency of Interest scores (*M* = ___, *SD* = ___, *n* = ___) compared to second-year students (*M* = ___, *SD* = ___, *n* = ___). For Perseverance of Effort, first-year students (*M* = ___, *SD* = ___) [similar/different] compared to second-year students (*M* = ___, *SD* = ___). These descriptive patterns suggest [your interpretation].
:::

### Your Write-Up

Write your paragraph here (or in your RStudio exercise):

::: {.callout-warning}
## Your Turn

Look at your summary statistics above and fill in the template. Think about:

- Which subscale shows larger differences between study years?
- Are first-year or second-year students higher on Consistency of Interest?
- What might explain differences in Perseverance of Effort?
:::

---

## Self-Assessment

Check your understanding before finishing.

::: {.webex-check .webex-box}

**1. In APA style, M stands for:**

`r mcq(c("Maximum", answer = "Mean", "Minimum", "Median"))`

**2. How many decimal places should you use for M and SD?**

`r mcq(c("0", "1", answer = "2", "3"))`

**3. True or False: You only need to report means, not standard deviations.**

`r torf(FALSE)`

**4. Which is correct APA format?**

`r mcq(c("Mean = 3.27", "m = 3.27", answer = "M = 3.27", "MEAN = 3.27"))`

**5. A good results paragraph includes M, SD, and `r fitb("n")` (sample size).**

:::

---

## Lab 1 Complete!

Congratulations! You've learned:

| Part | Skills |
|------|--------|
| 1 | R basics, downloading notebook |
| 2 | Loading data, inspecting structure |
| 3 | Reverse scoring Grit Scale items |
| 4 | Calculating subscales, grouped descriptives, boxplots |
| 5 | APA-style reporting |

---

## What's Next?

::: {.callout-tip}
## Coming Up in Lab 2

- **Independent samples t-test** - comparing two groups
- **Paired samples t-test** - comparing pre vs post
- **Effect sizes** - Cohen's d
- **APA reporting** for inferential statistics
:::

---

## Extra Practice

Want more practice? Try the follow-on exercise with a different dataset:

::: {.callout-important}
## Follow-On Exercise

The follow-on exercise is a **`.qmd` notebook file** that you download and run in **RStudio Server** on your own time. It uses a different dataset (Wellbeing Study) to give you extra practice applying the same skills you learned in Lab 1.

**To use it:**
1. Download the `.qmd` file below
2. Upload it to your RStudio Server workspace (same way you did with the main exercise)
3. Work through it independently to reinforce your learning
:::

```{r}
#| echo: false
downloadthis::download_file(
  path = "../../followon/lab1-followon.qmd",
  output_name = "lab1-followon",
  button_label = "Download Follow-On Exercise (.qmd)",
  button_type = "info",
  icon = "fa fa-download",
  class = "btn-lg"
)
```

---

## Navigation

::: {.d-flex .justify-content-between}
[← Back to Part 4](part4.qmd){.btn .btn-secondary}

[Back to Lab 1 Overview](index.qmd){.btn .btn-primary}
:::
