---
title: "Part 3: Reverse Scoring"
subtitle: "Lab 1: Descriptive Statistics"
format: html
filters:
  - webr
webr:
  packages: ['readr']
  show-startup-message: false
---

```{r, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Watch the Demo

Follow along with the slides below.

```{=html}
<iframe src="part3-slides.html" width="100%" height="500px" style="border: 1px solid #ccc; border-radius: 8px;"></iframe>
```

::: {.callout-tip}
## Tip
Click inside the slide area first, then use arrow keys to navigate.
:::

---

## Try It Yourself

::: {.callout-tip}
## How to Practice
1. **Run the code** to see how reverse scoring works
2. **Verify** the reversal by checking original + reversed = 6
3. **Copy to RStudio** and try with your own data
:::

### Load the Data First

```{webr-r}
library(readr)
data_url <- "https://raw.githubusercontent.com/LittleMonkeyLab/datarepo/main/personality_lab.csv"
data <- read_csv(data_url, show_col_types = FALSE)

cat("Data loaded! Items p4 and p5 need reversing.")
```

### Look at Original Values

```{webr-r}
library(readr)
data_url <- "https://raw.githubusercontent.com/LittleMonkeyLab/datarepo/main/personality_lab.csv"
data <- read_csv(data_url, show_col_types = FALSE)

# See the original p4 values
head(data$p4, 10)
```

### Apply the Reverse Formula

For a 1-5 scale: `reversed = 6 - original`

```{webr-r}
library(readr)
data_url <- "https://raw.githubusercontent.com/LittleMonkeyLab/datarepo/main/personality_lab.csv"
data <- read_csv(data_url, show_col_types = FALSE)

# Reverse score p4
data$rev_p4 <- 6 - data$p4

# Compare original and reversed
head(data[, c("p4", "rev_p4")])
```

::: {.callout-note}
## Challenge
What would the formula be for a 1-7 scale? Try it!
:::

### Verify the Reversal

Original + reversed should always equal 6:

```{webr-r}
library(readr)
data_url <- "https://raw.githubusercontent.com/LittleMonkeyLab/datarepo/main/personality_lab.csv"
data <- read_csv(data_url, show_col_types = FALSE)

data$rev_p4 <- 6 - data$p4

# Check: should all be 6
head(data$p4 + data$rev_p4)
```

### Reverse Both Items

Now reverse p5 as well:

```{webr-r}
library(readr)
data_url <- "https://raw.githubusercontent.com/LittleMonkeyLab/datarepo/main/personality_lab.csv"
data <- read_csv(data_url, show_col_types = FALSE)

# Reverse both items
data$rev_p4 <- 6 - data$p4
data$rev_p5 <- 6 - data$p5

# Check p5
head(data[, c("p5", "rev_p5")])
```

### Create Total Score

Sum all items using the **reversed** versions:

```{webr-r}
library(readr)
data_url <- "https://raw.githubusercontent.com/LittleMonkeyLab/datarepo/main/personality_lab.csv"
data <- read_csv(data_url, show_col_types = FALSE)

data$rev_p4 <- 6 - data$p4
data$rev_p5 <- 6 - data$p5

# Create personality total score
data$pers_score <- data$p1 + data$p2 + data$p3 +
                   data$rev_p4 + data$rev_p5

# View result
head(data[, c("id", "group", "time", "pers_score")])
```

---

## Self-Assessment

Check your understanding before moving on.

::: {.webex-check .webex-box}

**1. Why do we reverse-score some questionnaire items?**

`r mcq(c("To make the data look better", answer = "Because some items are negatively worded", "To reduce the sample size", "To increase variability"))`

**2. For a 1-5 scale, the reverse scoring formula is: `r fitb("6")` - original**

**3. If the original score is 2 on a 1-5 scale, the reversed score is `r fitb("4")`**

**4. True or False: When creating a total score, you should use the original (unreversed) items.**

`r torf(FALSE)`

**5. To verify reverse scoring worked, original + reversed should equal:**

`r mcq(c("0", "5", answer = "6 (for 1-5 scale)", "10"))`

:::

---

## Download Exercise

Continue practicing in RStudio:

```{r}
#| echo: false
downloadthis::download_file(
  path = "../../templates/lab1-exercise.qmd",
  output_name = "lab1-exercise",
  button_label = "Download Exercise (.qmd)",
  button_type = "warning",
  icon = "fa fa-download",
  class = "btn-lg"
)
```

---

## Next Steps

::: {.d-flex .justify-content-between}
[← Back to Part 2](part2.qmd){.btn .btn-secondary}

[Continue to Part 4: Descriptives & Plots →](part4.qmd){.btn .btn-primary}
:::
