---
title: "Part 5: APA Write-Up"
subtitle: "Lab 1: Descriptive Statistics"
format: html
filters:
  - webr
webr:
  packages: ['readr', 'dplyr']
  show-startup-message: false
---

```{r, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Watch the Demo

Follow along with the slides below.

```{=html}
<iframe src="part5-slides.html" width="100%" height="500px" style="border: 1px solid #ccc; border-radius: 8px;"></iframe>
```

::: {.callout-tip}
## Tip
Click inside the slide area first, then use arrow keys to navigate.
:::

---

## Try It Yourself

::: {.callout-tip}
## How to Practice
1. **Generate your summary statistics**
2. **Use the template** to write your paragraph
3. **Copy to your RStudio exercise** for submission
:::

### Generate Your Statistics

Run this to get all the numbers you need:

```{webr-r}
library(readr)
library(dplyr)

# Load and prepare data
data_url <- "https://raw.githubusercontent.com/LittleMonkeyLab/datarepo/main/personality_lab.csv"
data <- read_csv(data_url, show_col_types = FALSE)
data$rev_p4 <- 6 - data$p4
data$rev_p5 <- 6 - data$p5
data$pers_score <- data$p1 + data$p2 + data$p3 + data$rev_p4 + data$rev_p5

# Overall statistics
cat("=== OVERALL ===\n")
cat("M =", round(mean(data$pers_score), 2), "\n")
cat("SD =", round(sd(data$pers_score), 2), "\n")
cat("N =", length(data$pers_score), "\n\n")

# By group and time
cat("=== BY GROUP AND TIME ===\n")
summary_stats <- data |>
  group_by(group, time) |>
  summarise(
    M = round(mean(pers_score), 2),
    SD = round(sd(pers_score), 2),
    n = n(),
    .groups = "drop"
  )

print(summary_stats)
```

### Create a Formatted Table

```{webr-r}
library(readr)
library(dplyr)

data_url <- "https://raw.githubusercontent.com/LittleMonkeyLab/datarepo/main/personality_lab.csv"
data <- read_csv(data_url, show_col_types = FALSE)
data$rev_p4 <- 6 - data$p4
data$rev_p5 <- 6 - data$p5
data$pers_score <- data$p1 + data$p2 + data$p3 + data$rev_p4 + data$rev_p5

summary_stats <- data |>
  group_by(group, time) |>
  summarise(
    M = round(mean(pers_score), 2),
    SD = round(sd(pers_score), 2),
    n = n(),
    .groups = "drop"
  )

knitr::kable(summary_stats,
             caption = "Table 1. Descriptive Statistics by Group and Time",
             col.names = c("Group", "Time", "M", "SD", "n"))
```

---

## APA Write-Up Template

Use this template for your results paragraph. Fill in the blanks with your statistics:

::: {.callout-note}
## Template

Participants reported [describe level] levels of personality overall (*M* = ___, *SD* = ___). At baseline, the control group (*M* = ___, *SD* = ___, *n* = ___) and training group (*M* = ___, *SD* = ___, *n* = ___) showed [similar/different] scores. Following the intervention, the control group showed [description] (*M* = ___, *SD* = ___), whereas the training group showed [description] (*M* = ___, *SD* = ___). These descriptive patterns suggest [your interpretation].
:::

### Your Write-Up

Write your paragraph here (or in your RStudio exercise):

::: {.callout-warning}
## Your Turn

Look at your summary statistics above and fill in the template. Think about:

- Are the groups similar at baseline?
- Which group changed more from pre to post?
- What does this pattern suggest about the training?
:::

---

## Self-Assessment

Check your understanding before finishing.

::: {.webex-check .webex-box}

**1. In APA style, M stands for:**

`r mcq(c("Maximum", answer = "Mean", "Minimum", "Median"))`

**2. How many decimal places should you use for M and SD?**

`r mcq(c("0", "1", answer = "2", "3"))`

**3. True or False: You only need to report means, not standard deviations.**

`r torf(FALSE)`

**4. Which is correct APA format?**

`r mcq(c("Mean = 3.27", "m = 3.27", answer = "M = 3.27", "MEAN = 3.27"))`

**5. A good results paragraph includes M, SD, and `r fitb("n")` (sample size).**

:::

---

## Lab 1 Complete!

Congratulations! You've learned:

| Part | Skills |
|------|--------|
| 1 | R basics, variables, vectors |
| 2 | Loading data, inspecting structure |
| 3 | Reverse scoring questionnaires |
| 4 | Grouped descriptives, boxplots |
| 5 | APA-style reporting |

---

## Download Final Exercise

Complete and submit your exercise in RStudio:

```{r}
#| echo: false
downloadthis::download_file(
  path = "../../templates/lab1-exercise.qmd",
  output_name = "lab1-exercise",
  button_label = "Download Exercise (.qmd)",
  button_type = "warning",
  icon = "fa fa-download",
  class = "btn-lg"
)
```

---

## What's Next?

::: {.callout-tip}
## Coming Up in Lab 2

- **Independent samples t-test** - comparing two groups
- **Paired samples t-test** - comparing pre vs post
- **Effect sizes** - Cohen's d
- **APA reporting** for inferential statistics
:::

---

## Extra Practice

Want more practice? Try the follow-on exercise with a different dataset:

[Follow-On Exercise: Wellbeing Study](../../followon/lab1-followon.qmd){.btn .btn-secondary}

---

## Navigation

::: {.d-flex .justify-content-between}
[← Back to Part 4](part4.qmd){.btn .btn-secondary}

[Back to Lab 1 Overview](index.qmd){.btn .btn-primary}
:::
