---
title: "Part 1: What is R?"
subtitle: "Lab 1: Descriptive Statistics"
format: html
filters:
  - webr
webr:
  packages: []
  show-startup-message: false
---

```{r, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Watch the Demo

Follow along with the slides below. You can advance slides with the arrow keys or click the arrows.

```{=html}
<iframe src="part1-slides.html" width="100%" height="500px" style="border: 1px solid #ccc; border-radius: 8px;"></iframe>
```

::: {.callout-tip}
## Tip
Click inside the slide area first, then use arrow keys to navigate.
:::

---

## Try It Yourself

Now practice what you just learned! Run each code cell below.

::: {.callout-tip}
## How to Practice
1. **Run the code** by clicking the "Run Code" button
2. **Edit the code** - change numbers or add new lines, then run again
3. **Copy to RStudio** - try the same code in your RStudio Server console to see it work there too!
:::

### Basic Calculations

R works like a calculator. Try running this, then try changing the numbers:

```{webr-r}
1 + 1
```

What about more complex maths?

```{webr-r}
10 * 5 + 2
```

### Creating Variables

Use the assignment operator `<-` to store values. Try changing the values and see what happens:

```{webr-r}
x <- 5
y <- 3
x + y
```

::: {.callout-note}
## Challenge
Can you create a variable called `age` and set it to your age? Then create `next_year` that adds 1 to it?
:::

### Creating Vectors

A vector holds multiple values. Create one with `c()`. Try adding more numbers:

```{webr-r}
scores <- c(4, 5, 3, 2, 5)
scores
```

### Using Functions

Calculate descriptive statistics. Copy these to RStudio and try with your own data:

```{webr-r}
scores <- c(4, 5, 3, 2, 5)
mean(scores)
```

```{webr-r}
scores <- c(4, 5, 3, 2, 5)
sd(scores)
```

```{webr-r}
scores <- c(4, 5, 3, 2, 5)
length(scores)
```

---

## Self-Assessment

Check your understanding before moving on.

::: {.webex-check .webex-box}

**1. What symbol is used to assign a value to a variable in R?**

`r mcq(c("=", answer = "<-", "->", "=="))`

**2. What function creates a vector of multiple values?**

`r mcq(c("vector()", "v()", answer = "c()", "combine()"))`

**3. The mean of `c(2, 4, 6)` is `r fitb("4")`.**

**4. True or False: R is free and open source.**

`r torf(TRUE)`

**5. True or False: RStudio and R are the same thing.**

`r torf(FALSE)`

:::

---

## Download Exercise

Ready to practice in RStudio? Download the exercise file and open it in RStudio Server:

```{r}
#| echo: false
downloadthis::download_file(
  path = "../../templates/lab1-exercise.qmd",
  output_name = "lab1-exercise",
  button_label = "Download Exercise (.qmd)",
  button_type = "warning",
  icon = "fa fa-download",
  class = "btn-lg"
)
```

---

## Next Steps

::: {.d-flex .justify-content-between}
[← Back to Lab 1 Overview](../../labs/lab01.qmd){.btn .btn-secondary}

[Continue to Part 2: Loading Data →](part2.qmd){.btn .btn-primary}
:::
