---
title: "Lab 2: Comparing Groups"
author: "Your Name"
date: "`r format(Sys.Date(), '%d %B %Y')`"
format:
  html:
    toc: true
    code-fold: false
    code-annotations: below
    theme: cosmo
---

## 1. Load Packages

```{r}
#| message: false
#| warning: false

library(dplyr)
library(ggplot2)
library(readr)
```

---

## 2. Load and Inspect Data

```{r}
# TODO: Load your dataset
# data <- read_csv("your_data.csv")

# Check group sizes
# data |> count(group)
```

---

## 3. Descriptive Statistics by Group

```{r}
# TODO: Compute descriptives for each group
# data |>
#   group_by(group) |>
#   summarise(
#     mean = mean(score, na.rm = TRUE),
#     sd = sd(score, na.rm = TRUE),
#     n = n()
#   )
```

---

## 4. Independent Samples t-test

```{r}
# TODO: Run independent samples t-test
# t.test(score ~ group, data = data)
```

### Effect Size (Cohen's d)

```{r}
# TODO: Calculate Cohen's d
# group1 <- data |> filter(group == "A") |> pull(score)
# group2 <- data |> filter(group == "B") |> pull(score)
#
# mean_diff <- mean(group2) - mean(group1)
# pooled_sd <- sqrt((sd(group1)^2 + sd(group2)^2) / 2)
# cohens_d <- mean_diff / pooled_sd
# cohens_d
```

---

## 5. Paired Samples t-test (if applicable)

```{r}
# TODO: Run paired t-test (for repeated measures)
# t.test(data$pre, data$post, paired = TRUE)
```

---

## 6. Visualisation

```{r}
# TODO: Create comparison plot
# ggplot(data, aes(x = group, y = score, fill = group)) +
#   geom_boxplot() +
#   geom_jitter(width = 0.1, alpha = 0.5) +
#   theme_minimal() +
#   labs(title = "Comparison of Groups")
```

---

## 7. APA Write-Up

[Write your results in APA format. Example:]

> An independent samples t-test revealed a significant difference between the control group (*M* = ___, *SD* = ___) and treatment group (*M* = ___, *SD* = ___), *t*(___) = ___, *p* = ___, *d* = ___.

---

## 8. Reflection

1. Was the difference statistically significant?

   [Your answer]

2. Was the effect size small, medium, or large?

   [Your answer]

3. What does this mean practically?

   [Your answer]
