---
title: "Lab 3: ANOVA"
author: "Your Name"
date: "`r format(Sys.Date(), '%d %B %Y')`"
format:
  html:
    toc: true
    code-fold: false
    code-annotations: below
    theme: cosmo
---

## 1. Load Packages

```{r}
#| message: false
#| warning: false

library(dplyr)
library(ggplot2)
library(readr)
```

---

## 2. Load and Inspect Data

```{r}
# TODO: Load your dataset
# data <- read_csv("your_data.csv")

# Check group sizes
# data |> count(condition)
```

---

## 3. Descriptive Statistics by Group

```{r}
# TODO: Compute descriptives for each group
# data |>
#   group_by(condition) |>
#   summarise(
#     mean = mean(score, na.rm = TRUE),
#     sd = sd(score, na.rm = TRUE),
#     n = n()
#   )
```

---

## 4. One-Way ANOVA

```{r}
# TODO: Fit the ANOVA model
# model <- aov(score ~ condition, data = data)

# View results
# summary(model)
```

---

## 5. Post-Hoc Tests

```{r}
# TODO: Run Tukey's HSD (if ANOVA is significant)
# TukeyHSD(model)
```

---

## 6. Effect Size (Eta-Squared)

```{r}
# TODO: Calculate eta-squared
# model_summary <- summary(model)
# ss_between <- model_summary[[1]]["condition", "Sum Sq"]
# ss_residual <- model_summary[[1]]["Residuals", "Sum Sq"]
# eta_sq <- ss_between / (ss_between + ss_residual)
# eta_sq
```

---

## 7. Visualisation

```{r}
# TODO: Create boxplot by group
# ggplot(data, aes(x = condition, y = score, fill = condition)) +
#   geom_boxplot(alpha = 0.7) +
#   geom_jitter(width = 0.1, alpha = 0.5) +
#   theme_minimal() +
#   labs(title = "Scores by Condition")
```

---

## 8. Two-Way ANOVA (if applicable)

```{r}
# TODO: Fit factorial ANOVA (if you have two factors)
# model2 <- aov(score ~ factor_a * factor_b, data = data)
# summary(model2)
```

---

## 9. APA Write-Up

[Write your results in APA format. Example:]

> A one-way ANOVA revealed a significant effect of condition on scores, *F*(2, 97) = ___, *p* = ___, η² = ___.

> Post-hoc comparisons using Tukey's HSD indicated that...

---

## 10. Reflection

1. Was the overall ANOVA significant?

   [Your answer]

2. Which groups differed from each other (based on post-hoc tests)?

   [Your answer]

3. How would you interpret the effect size?

   [Your answer]
