---
title: "Lab 4: Linear Models & Correlation"
author: "Your Name"
date: "`r format(Sys.Date(), '%d %B %Y')`"
format:
  html:
    toc: true
    code-fold: false
    code-annotations: below
    theme: cosmo
---

## 1. Load Packages

```{r}
#| message: false
#| warning: false

library(dplyr)
library(ggplot2)
library(readr)
```

---

## 2. Load and Inspect Data

```{r}
# TODO: Load your dataset
# data <- read_csv("your_data.csv")

# Check structure
# str(data)
```

---

## 3. Correlation Analysis

### 3a. Correlation test

```{r}
# TODO: Test correlation between two variables
# cor.test(data$var1, data$var2)
```

### 3b. Scatterplot

```{r}
# TODO: Visualise the relationship
# ggplot(data, aes(x = var1, y = var2)) +
#   geom_point() +
#   geom_smooth(method = "lm", se = TRUE) +
#   theme_minimal() +
#   labs(title = "Relationship between X and Y")
```

---

## 4. Simple Linear Regression

```{r}
# TODO: Build regression model
# model <- lm(outcome ~ predictor, data = data)
# summary(model)
```

### Interpretation

- **Intercept**: [Interpret]
- **Slope**: [Interpret]
- **R-squared**: [Interpret]
- **p-value**: [Interpret]

---

## 5. Multiple Regression

```{r}
# TODO: Add multiple predictors
# model2 <- lm(outcome ~ pred1 + pred2 + pred3, data = data)
# summary(model2)
```

---

## 6. Model Diagnostics

```{r}
# TODO: Check assumptions
# par(mfrow = c(2, 2))
# plot(model)
```

### Assumptions checklist

- [ ] Linearity
- [ ] Normality of residuals
- [ ] Homoscedasticity
- [ ] Independence

---

## 7. APA Write-Up

[Write your results in APA format. Example:]

> A simple linear regression was conducted to predict [outcome] from [predictor]. The model was statistically significant, *F*(1, ___) = ___, *p* < .001, *R*² = ___. [Predictor] significantly predicted [outcome], *b* = ___, *t*(___) = ___, *p* = ___.

---

## 8. Reflection

1. How strong was the relationship?

   [Your answer]

2. How much variance did your model explain?

   [Your answer]

3. Were all assumptions met?

   [Your answer]

---

## Congratulations!

You've completed all 4 labs of the GoldPsychR intensive programme!
