---
title: "Lab 1: Descriptive Statistics"
subtitle: "Grit & The Short Grit Scale"
---

# Integrated Lab Experience

::: {.callout-tip}
## Recommended Workflow

Work through the integrated lab experience below. You only need two browser tabs: this page and RStudio Server.
:::

## Start the Interactive Lab

[**Start Lab 1**](../chapters/lab1/index.qmd){.btn .btn-success .btn-lg}

---

## Quick Navigation to Parts

If you need to jump back to a specific part:

:::: {.columns}

::: {.column width="50%"}
### ![](/assets/images/svg/laptop.svg){width=20} [Part 1: What is R?](../chapters/lab1/part1.qmd)
- RStudio orientation
- Variables and vectors
- Basic functions
- **Download notebook**
:::

::: {.column width="50%"}
### ![](/assets/images/svg/table.svg){width=20} [Part 2: Loading Data](../chapters/lab1/part2.qmd)
- Reading CSV files
- Exploring data structure
- Understanding data types
:::

::::

:::: {.columns}

::: {.column width="50%"}
### ![](/assets/images/svg/code.svg){width=20} [Part 3: Reverse Scoring](../chapters/lab1/part3.qmd)
- Identifying reverse-scored items
- Calculating reverse scores
- Creating subscale scores
:::

::: {.column width="50%"}
### ![](/assets/images/svg/chart.svg){width=20} [Part 4: Descriptives & Plots](../chapters/lab1/part4.qmd)
- Group summaries with dplyr
- Boxplots with ggplot2
- Interpreting visualisations
:::

::::

:::: {.columns}

::: {.column width="50%"}
### ![](/assets/images/svg/document.svg){width=20} [Part 5: APA Write-Up](../chapters/lab1/part5.qmd)
- Reporting M, SD, n
- Describing group differences
- Professional formatting
:::

::: {.column width="50%"}
### ![](/assets/images/svg/rocket.svg){width=20} Extra Practice
[Follow-On Exercise](../followon/lab1-followon.qmd) - Apply skills to a new dataset
:::

::::

---

# Overview

In this lab, you'll learn to analyse Grit Scale data in R. We'll work through loading data, reverse-scoring questionnaire items, calculating subscales, creating visualisations, and writing up results.

## The Data Story

We've collected data using the Short Grit Scale (Grit-S):

- **120 university students** completed the 8-item Grit Scale
- Participants are either **First Year** or **Final Year** students
- The scale measures two subscales: Consistency of Interest and Perseverance of Effort

**Our research question**: Do grit scores differ between first-year and final-year students?

::: {.callout-note}
![](/assets/images/svg/brain.svg){width=20 style="float: left; margin-right: 10px;"}

## Learn More About the Grit Scale

Want to read the original research? [Access the source article](https://librarysearch.gold.ac.uk/permalink/f/igqbsh/TN_cdi_informaworld_taylorfrancis_310_1080_00223890802634290) through the Goldsmiths Library.
:::

## ![](/assets/images/svg/target.svg){width=24} Learning Objectives

By the end of this lab, you will be able to:

- Navigate RStudio and understand its four panes
- Load CSV data into R using `read_csv()`
- Reverse-score negatively worded questionnaire items (4 items: G1, G3, G5, G6)
- Calculate subscale scores (Consistency of Interest, Perseverance of Effort)
- Calculate descriptive statistics by group using `dplyr`
- Create boxplots using `ggplot2`
- Write an APA-style results paragraph

# Schedule

**Lab Time: 13:00 - 15:00**

| Time | Activity |
|------|----------|
| 13:00-13:10 | Part 1: What is R? + RStudio Tour |
| 13:10-13:20 | **Practice: WebR basics** |
| 13:20-13:35 | Part 2: Loading Data |
| 13:35-13:50 | **Practice: Data loading** |
| 13:50-14:05 | Part 3: Reverse Scoring |
| 14:05-14:20 | **Practice: Scoring** |
| 14:20-14:40 | Part 4: Descriptives & Visualisation |
| 14:40-14:55 | **Practice: Analysis + Write-Up** |
| 14:55-15:00 | Wrap-up |

# Key Concepts

## The Five Parts

| Part | Topic | Key Skills |
|------|-------|------------|
| 1 | RStudio Tour | Variables, vectors, functions, **download notebook** |
| 2 | Loading Data | `read_csv()`, `head()`, `str()`, `dim()` |
| 3 | Reverse Scoring | Formula: `6 - original` (4 items: G1, G3, G5, G6) |
| 4 | Descriptives & Plots | `group_by()`, `summarise()`, `ggplot()` boxplots |
| 5 | APA Write-Up | Reporting M, SD, n, proactive coding |

---

## Alternative Resources

Or access individual resources separately:

### Instructor Slides

![](/assets/images/svg/slides.svg){width=20} [**Open Full Slides**](../slides/lab1-slides.qmd){.btn .btn-primary target="_blank"}

### Browser Practice

![](/assets/images/svg/play.svg){width=20} [**Open WebR Practice**](../chapters/lab1-webr.qmd){.btn .btn-outline-success}

### RStudio Exercise

![](/assets/images/svg/laptop.svg){width=20} Download the exercise notebook to complete in RStudio Server:

```{r}
#| echo: false
downloadthis::download_file(
  path = "../templates/lab1-exercise.qmd",
  output_name = "lab1-exercise",
  button_label = "Download Exercise (.qmd)",
  button_type = "warning",
  icon = "fa fa-download",
  class = "btn-lg"
)
```

### Extra Practice

![](/assets/images/svg/document.svg){width=20} Download the follow-on exercise for independent practice:

```{r}
#| echo: false
downloadthis::download_file(
  path = "../followon/lab1-followon.qmd",
  output_name = "lab1-followon",
  button_label = "Download Follow-On Exercise (.qmd)",
  button_type = "info",
  icon = "fa fa-download",
  class = "btn-lg"
)
```

# Help

::: {.callout-note}
## Need Help?

- **During the lab**: Ask your instructor or demonstrators
- **Reference**: Check the [R Toolbox](../toolbox/index.qmd) for code examples
- **Errors**: See the [Troubleshooting Guide](../toolbox/troubleshooting.qmd)
:::
