---
title: "Lab 2: T-Tests"
subtitle: "Test Anxiety & Statistical Inference"
---

# Integrated Lab Experience

::: {.callout-tip}
## Recommended Workflow

Work through the integrated lab experience below. You only need two browser tabs: this page and RStudio Server.
:::

### RStudio Exercise

![](/assets/images/svg/laptop.svg){width=20} Download the exercise notebook to complete in RStudio Server:

```{r}
#| echo: false
downloadthis::download_file(
  path = "../templates/lab2-exercise.qmd",
  output_name = "lab2-exercise",
  button_label = "Download Exercise (.qmd)",
  button_type = "warning",
  icon = "fa fa-download",
  class = "btn-lg"
)
```


## Start the Interactive Lab

[**Start Lab 2**](../chapters/lab2/index.qmd){.btn .btn-success .btn-lg}

---

## Quick Navigation to Parts

If you need to jump back to a specific part:

:::: {.columns}

::: {.column width="50%"}
### ![](/assets/images/svg/brain.svg){width=20} [Part 1: Introduction to T-Tests](../chapters/lab2/part1.qmd)
- Hypothesis testing concepts
- Types of t-tests
- When to use each test
- **Download notebook**
:::

::: {.column width="50%"}
### ![](/assets/images/svg/chart.svg){width=20} [Part 2: One-Sample T-Test](../chapters/lab2/part2.qmd)
- Comparing to a known value
- Test assumptions
- Interpreting output
:::

::::

:::: {.columns}

::: {.column width="50%"}
### ![](/assets/images/svg/group.svg){width=20} [Part 3: Independent Samples T-Test](../chapters/lab2/part3.qmd)
- Comparing two groups
- Levene's test for equality
- Welch's correction
:::

::: {.column width="50%"}
### ![](/assets/images/svg/code.svg){width=20} [Part 4: Paired Samples T-Test](../chapters/lab2/part4.qmd)
- Pre-post comparisons
- Within-subjects design
- Difference scores
:::

::::

:::: {.columns}

::: {.column width="50%"}
### ![](/assets/images/svg/document.svg){width=20} [Part 5: Effect Sizes & Reporting](../chapters/lab2/part5.qmd)
- Cohen's d calculation
- APA-style results
- Professional write-up
:::

::: {.column width="50%"}
### ![](/assets/images/svg/rocket.svg){width=20} Extra Practice
[Follow-On Exercise](../followon/lab2-followon.qmd) - Apply skills to a new dataset
:::

::::

---

# Overview

In this lab, you'll learn to perform and interpret t-tests in R. We'll work through hypothesis testing, running one-sample, independent samples, and paired samples t-tests, calculating effect sizes, and writing up results in APA format.

## The Data Story

We've collected data from a Test Anxiety intervention study:

- **80 university students** completed the Test Anxiety Inventory
- Participants were assigned to either **Control** or **Intervention** groups
- Anxiety was measured **before and after** the intervention
- The population mean for this scale is known to be **50**

**Our research questions**:
1. Does our sample differ from the population norm?
2. Do the groups differ at baseline?
3. Did anxiety change from pre to post?

::: {.callout-note}
![](/assets/images/svg/brain.svg){width=20 style="float: left; margin-right: 10px;"}

## Learn More About Test Anxiety

Test anxiety is a significant concern in educational settings. [Learn about its measurement and impact](https://librarysearch.gold.ac.uk/primo-explore/search?query=any,contains,test%20anxiety&tab=default_tab&search_scope=default_scope&vid=44GOL_VU1&lang=en_US&offset=0) through the Goldsmiths Library.
:::

## ![](/assets/images/svg/target.svg){width=24} Learning Objectives

By the end of this lab, you will be able to:

- Explain the logic of null hypothesis significance testing
- Choose the appropriate t-test for different research designs
- Conduct a one-sample t-test using `t.test()`
- Conduct an independent samples t-test comparing two groups
- Conduct a paired samples t-test for pre-post designs
- Calculate and interpret Cohen's d effect sizes
- Check assumptions for t-tests
- Write APA-style results for t-tests

# Schedule

**Lab Time: 13:00 - 15:00**

| Time | Activity |
|------|----------|
| 13:00-13:15 | Part 1: Introduction to T-Tests |
| 13:15-13:30 | **Practice: Hypothesis concepts** |
| 13:30-13:45 | Part 2: One-Sample T-Test |
| 13:45-14:00 | **Practice: One-sample analysis** |
| 14:00-14:15 | Part 3: Independent Samples T-Test |
| 14:15-14:30 | **Practice: Independent samples analysis** |
| 14:30-14:45 | Part 4: Paired Samples T-Test |
| 14:45-14:55 | Part 5: Effect Sizes & Write-Up |
| 14:55-15:00 | Wrap-up |

# Key Concepts

## The Five Parts

| Part | Topic | Key Skills |
|------|-------|------------|
| 1 | Introduction | Hypotheses, test selection, **download notebook** |
| 2 | One-Sample T-Test | `t.test(x, mu = )`, comparing to population |
| 3 | Independent Samples | `t.test(y ~ group)`, Welch's test |
| 4 | Paired Samples | `t.test(pre, post, paired = TRUE)` |
| 5 | Effect Sizes & Reporting | Cohen's d, APA format, interpretation |

---

## Alternative Resources

Or access individual resources separately:

### Instructor Slides

![](/assets/images/svg/slides.svg){width=20} [**Open Full Slides**](../slides/lab2-slides.qmd){.btn .btn-primary target="_blank"}

### Browser Practice

![](/assets/images/svg/play.svg){width=20} [**Open WebR Practice**](../chapters/lab2-webr.qmd){.btn .btn-outline-success}

### RStudio Exercise

![](/assets/images/svg/laptop.svg){width=20} Download the exercise notebook to complete in RStudio Server:

```{r}
#| echo: false
downloadthis::download_file(
  path = "../templates/lab2-exercise.qmd",
  output_name = "lab2-exercise",
  button_label = "Download Exercise (.qmd)",
  button_type = "warning",
  icon = "fa fa-download",
  class = "btn-lg"
)
```

### Extra Practice

![](/assets/images/svg/document.svg){width=20} Download the follow-on exercise for independent practice:

```{r}
#| echo: false
downloadthis::download_file(
  path = "../followon/lab2-followon.qmd",
  output_name = "lab2-followon",
  button_label = "Download Follow-On Exercise (.qmd)",
  button_type = "info",
  icon = "fa fa-download",
  class = "btn-lg"
)
```

# Help

::: {.callout-note}
## Need Help?

- **During the lab**: Ask your instructor
- **Reference**: Check the [R Toolbox](../toolbox/index.qmd) for code examples
- **Errors**: See the [Troubleshooting Guide](../toolbox/troubleshooting.qmd)
:::
