---
title: "Lab 3: ANOVA"
subtitle: "Study Techniques & Comparing Multiple Groups"
---

# Integrated Lab Experience

::: {.callout-tip}
## Recommended Workflow

Work through the integrated lab experience below. You only need two browser tabs: this page and RStudio Server.
:::

## Start the Interactive Lab

[**Start Lab 3**](../chapters/lab3/index.qmd){.btn .btn-success .btn-lg}

---

## Quick Navigation to Parts

If you need to jump back to a specific part:

:::: {.columns}

::: {.column width="50%"}
### ![](/assets/images/svg/brain.svg){width=20} [Part 1: Introduction to ANOVA](../chapters/lab3/part1.qmd)
- When t-tests aren't enough
- The logic of ANOVA
- F-ratio explained
- **Download notebook**
:::

::: {.column width="50%"}
### ![](/assets/images/svg/chart.svg){width=20} [Part 2: One-Way ANOVA](../chapters/lab3/part2.qmd)
- Comparing 3+ groups
- Using `aov()` function
- Interpreting output
:::

::::

:::: {.columns}

::: {.column width="50%"}
### ![](/assets/images/svg/target.svg){width=20} [Part 3: Post-Hoc Tests](../chapters/lab3/part3.qmd)
- Tukey's HSD
- Pairwise comparisons
- Controlling Type I error
:::

::: {.column width="50%"}
### ![](/assets/images/svg/group.svg){width=20} [Part 4: Two-Way ANOVA](../chapters/lab3/part4.qmd)
- Factorial designs
- Main effects
- Interaction effects
:::

::::

:::: {.columns}

::: {.column width="50%"}
### ![](/assets/images/svg/document.svg){width=20} [Part 5: Effect Sizes & Reporting](../chapters/lab3/part5.qmd)
- Eta-squared (η²)
- Partial eta-squared
- APA-style results
:::

::: {.column width="50%"}
### ![](/assets/images/svg/rocket.svg){width=20} Extra Practice
[Follow-On Exercise](../followon/lab3-followon.qmd) - Apply skills to a new dataset
:::

::::

---

# Overview

In this lab, you'll learn to compare three or more groups using Analysis of Variance (ANOVA). We'll work through one-way and two-way designs, conduct post-hoc comparisons, calculate effect sizes, and write up results in APA format.

## The Data Story

We've collected data from a Study Techniques experiment:

- **120 university students** participated in a memory test
- Students used one of **three study methods**: Rereading, Self-Testing, or Elaboration
- Half studied in the **Morning**, half in the **Evening**
- Outcome: Number of items correctly recalled (0-30)

**Our research questions**:
1. Do the three study methods differ in effectiveness?
2. Does time of day affect memory performance?
3. Is there an interaction between study method and time of day?

::: {.callout-note}
![](/assets/images/svg/brain.svg){width=20 style="float: left; margin-right: 10px;"}

## Learn More About Study Techniques

The testing effect is one of the most robust findings in cognitive psychology. [Explore the research](https://librarysearch.gold.ac.uk/discovery/search?query=any,contains,testing%20effect%20memory&tab=Everything&search_scope=MyInst_and_CI&vid=44GOL_INST:44GOL_VU1&offset=0) through the Goldsmiths Library.
:::

## ![](/assets/images/svg/target.svg){width=24} Learning Objectives

By the end of this lab, you will be able to:

- Explain when to use ANOVA instead of t-tests
- Conduct a one-way ANOVA using `aov()`
- Interpret F-ratios and p-values from ANOVA output
- Conduct Tukey's HSD post-hoc comparisons
- Conduct a two-way (factorial) ANOVA
- Identify and interpret main effects and interactions
- Calculate and interpret eta-squared (η²)
- Write APA-style ANOVA results

# Schedule

**Lab Time: 13:00 - 15:00**

| Time | Activity |
|------|----------|
| 13:00-13:15 | Part 1: Introduction to ANOVA |
| 13:15-13:30 | **Practice: ANOVA concepts** |
| 13:30-13:45 | Part 2: One-Way ANOVA |
| 13:45-14:00 | **Practice: One-way analysis** |
| 14:00-14:15 | Part 3: Post-Hoc Tests |
| 14:15-14:30 | **Practice: Tukey's HSD** |
| 14:30-14:45 | Part 4: Two-Way ANOVA |
| 14:45-14:55 | Part 5: Effect Sizes & Write-Up |
| 14:55-15:00 | Wrap-up |

# Key Concepts

## The Five Parts

| Part | Topic | Key Skills |
|------|-------|------------|
| 1 | Introduction | F-ratio, between/within variance, **download notebook** |
| 2 | One-Way ANOVA | `aov(y ~ group)`, `summary()` |
| 3 | Post-Hoc Tests | `TukeyHSD()`, pairwise comparisons |
| 4 | Two-Way ANOVA | `aov(y ~ A * B)`, interactions |
| 5 | Effect Sizes & Reporting | η², partial η², APA format |

---

## Alternative Resources

Or access individual resources separately:

### Instructor Slides

![](/assets/images/svg/slides.svg){width=20} [**Open Full Slides**](../slides/lab3-slides.qmd){.btn .btn-primary target="_blank"}

### Browser Practice

![](/assets/images/svg/play.svg){width=20} [**Open WebR Practice**](../chapters/lab3-webr.qmd){.btn .btn-outline-success}

### RStudio Exercise

![](/assets/images/svg/laptop.svg){width=20} Download the exercise notebook to complete in RStudio Server:

```{r}
#| echo: false
downloadthis::download_file(
  path = "../templates/lab3-exercise.qmd",
  output_name = "lab3-exercise",
  button_label = "Download Exercise (.qmd)",
  button_type = "warning",
  icon = "fa fa-download",
  class = "btn-lg"
)
```

### Extra Practice

![](/assets/images/svg/document.svg){width=20} Download the follow-on exercise for independent practice:

```{r}
#| echo: false
downloadthis::download_file(
  path = "../followon/lab3-followon.qmd",
  output_name = "lab3-followon",
  button_label = "Download Follow-On Exercise (.qmd)",
  button_type = "info",
  icon = "fa fa-download",
  class = "btn-lg"
)
```

# Help

::: {.callout-note}
## Need Help?

- **During the lab**: Ask your instructor
- **Reference**: Check the [R Toolbox](../toolbox/index.qmd) for code examples
- **Errors**: See the [Troubleshooting Guide](../toolbox/troubleshooting.qmd)
:::
