---
title: "Lab 1: Descriptive Statistics"
subtitle: "Grit & The Short Grit Scale"
author: "YOUR NAME HERE"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    code-fold: false
    code-annotations: below
    theme: cosmo
    embed-resources: true
---

```{r}
#| label: setup
#| message: false
#| warning: false

# Load required packages
library(readr)
library(dplyr)
library(ggplot2)
```

# Introduction

In this lab, you will analyse Grit Scale data to practice:

- Loading and inspecting data
- Reverse-scoring questionnaire items
- Calculating subscale scores
- Calculating descriptive statistics
- Creating visualisations
- Writing APA-style results

::: {.callout-important}
## How This Lab Works: Watch → Try → Do

Every part follows the same **3-step cycle**:

1. **Watch** - Follow along with the instructor slides
2. **Try** - Practice with interactive code on the activity page
3. **Do** - Complete exercises in this notebook (that's this file!)

You'll work through this notebook across all 5 parts. Let's start with your first task!
:::

---

# Your First Task: Change Your Name and Render

::: {.callout-note}
## Slides Part 1
Complete this task during **Part 1: What is R?** practice time.
:::

## Step 1: Find Your Name

Look at the **very top** of this file (lines 1-5). You'll see:

```yaml
---
title: "Lab 1: Descriptive Statistics"
subtitle: "Grit & The Short Grit Scale"
author: "YOUR NAME HERE"
date: today
```

**Find line 4** where it says `author: "YOUR NAME HERE"`

## Step 2: Replace with Your Name

1. **Click** on `"YOUR NAME HERE"` 
2. **Type** your actual name (keep the quotation marks!)
3. For example: `author: "Jane Smith"` or `author: "Alex Johnson"`

## Step 3: Render Your Notebook

1. **Look at the top toolbar** in RStudio Server
2. **Find the Render button** - it looks like a knitting needle 🧶
3. **Click the Render button**
4. Wait a few seconds... RStudio will process your file
5. When it's done, a **preview window** will open showing your rendered HTML document
6. **Check the top** - you should see your name in the document header!

::: {.callout-success}
## Success!

If you see your name in the rendered HTML document, **congratulations!** You've just:
- ✅ Edited a Quarto document
- ✅ Rendered it to HTML
- ✅ Created your first output file

This is exactly how you'll work throughout the lab. Make changes → Render → See results!
:::

**What happened?** The Render button converted your `.qmd` file into a beautiful HTML document. Every time you make changes and render, you create a new version with your latest work. This is how you'll submit your assignments!

---

# Part 1: Setup and Overview

::: {.callout-note}
## Slides Part 1
Continue with this section during **Part 1: What is R?** practice time.
:::

## About the Grit Scale

The Short Grit Scale (Grit-S) measures trait-level perseverance and passion for long-term goals. It has 8 items that form two subscales:

- **Consistency of Interest** (4 items: G1, G3, G5, G6) - measures ability to maintain interest
- **Perseverance of Effort** (4 items: G2, G4, G7, G8) - measures ability to work hard despite setbacks

**Important**: Items G1, G3, G5, and G6 are reverse-scored.

::: {.callout-note}
## Source Article

Want to learn more about the Grit Scale? [Read the original research article](https://librarysearch.gold.ac.uk/permalink/f/igqbsh/TN_cdi_informaworld_taylorfrancis_310_1080_00223890802634290) through the Goldsmiths Library.
:::

::: {.callout-note}
## Exercise 1.1: Comprehension Check

Before we dive into the analysis, let's check your understanding:

1. **What are the two subscales of the Grit Scale?** 
   
   *Hint: Look at the "About the Grit Scale" section above. The subscales measure different aspects of grit.*
   
   **[YOUR ANSWER]**

2. **Which items need to be reverse-scored and why?** 
   
   *Hint: Check which items belong to the Consistency of Interest subscale. Why might these need to be reversed?* You can always look at the journal article to see if the answer is there.
   
   **[YOUR ANSWER]**

3. **What is our main research question for this lab?** 
   
   *Hint: Think about what we're comparing. Look at the dataset description - what are the two groups we're comparing?*
   
   **[YOUR ANSWER]**
:::

---

# Part 2: Load and Inspect Data

::: {.callout-note}
## Slides Part 2
Complete this section during **Part 2: Loading Data** practice time.
:::

## Load the Dataset

```{r}
#| label: load-data

# Load the Grit Scale data from GitHub
data_url <- "https://raw.githubusercontent.com/LittleMonkeyLab/datarepo/main/Y1_Week13_Lab1_grit_data.csv"
grit_data <- read_csv(data_url, show_col_types = FALSE)
```

## Inspect the Data

```{r}
#| label: inspect-head

# View the first 6 rows
head(grit_data)
```

```{r}
#| label: inspect-structure

# Check the structure
str(grit_data)
```

```{r}
#| label: inspect-dimensions

# Check dimensions
dim(grit_data)
```

```{r}
#| label: inspect-summary

# Get summary statistics
summary(grit_data)
```

::: {.callout-note}
## Exercise 2.1

Answer the following questions based on your inspection:

1. How many participants are in the dataset? **[YOUR ANSWER]**
2. How many variables are there? **[YOUR ANSWER]**
3. What are the two levels of the `Study_Year` variable? **[YOUR ANSWER]**
4. What is the range of values for the Grit items (G1-G8)? **[YOUR ANSWER]**
:::

---

# Part 3: Reverse Scoring

::: {.callout-note}
## Slides Part 3
Complete this section during **Part 3: Reverse Scoring** practice time.
:::

## Understanding Reverse-Scored Items

Items G1, G3, G5, and G6 are reverse-scored. These items measure lack of consistency, so we reverse them to align with the other items.

For a 1-5 Likert scale, we use the formula: `reversed = 6 - original`

## Create Reversed Items

```{r}
#| label: reverse-score
#| code-annotations: above

# Reverse-score items G1, G3, G5, G6
#| 1. Apply the reverse formula to G1
grit_data$rev_G1 <- 6 - grit_data$G1
#| 2. Apply the reverse formula to G3
grit_data$rev_G3 <- 6 - grit_data$G3
#| 3. Apply the reverse formula to G5
grit_data$rev_G5 <- 6 - grit_data$G5
#| 4. Apply the reverse formula to G6
grit_data$rev_G6 <- 6 - grit_data$G6

# Verify the reversal worked correctly
head(grit_data[, c("G1", "rev_G1", "G3", "rev_G3", "G5", "rev_G5", "G6", "rev_G6")])
```

::: {.callout-note}
## Exercise 3.1

Look at the output above. For the first participant:

- Original G1 score: **[YOUR ANSWER]**
- Reversed G1 score: **[YOUR ANSWER]**
- Do they sum to 6? **[YOUR ANSWER]**
:::

## Calculate Subscale Scores

```{r}
#| label: calculate-subscales
#| code-annotations: above

# Consistency of Interest subscale (use REVERSED versions of G1, G3, G5, G6)
#| 1. Calculate Consistency of Interest as mean of reversed G1, G3, G5, G6
grit_data$consistency_interest <- (grit_data$rev_G1 + grit_data$rev_G3 + 
                                    grit_data$rev_G5 + grit_data$rev_G6) / 4

# Perseverance of Effort subscale (forward-scored items G2, G4, G7, G8)
#| 2. Calculate Perseverance of Effort as mean of G2, G4, G7, G8
grit_data$perseverance_effort <- (grit_data$G2 + grit_data$G4 + 
                                  grit_data$G7 + grit_data$G8) / 4

# Total Grit score (average of all 8 items after reverse-scoring)
#| 3. Calculate Total Grit as mean of all 8 items (after reverse-scoring)
grit_data$total_grit <- (grit_data$rev_G1 + grit_data$rev_G3 + grit_data$rev_G5 + grit_data$rev_G6 +
                         grit_data$G2 + grit_data$G4 + grit_data$G7 + grit_data$G8) / 8

# View the new columns
head(grit_data[, c("Study_Year", "consistency_interest", "perseverance_effort", "total_grit")])
```

::: {.callout-note}
## Exercise 3.2

Check your calculations:

- What is the range of possible scores for each subscale? **[YOUR ANSWER]**
- What is the range for total grit? **[YOUR ANSWER]**
- Do the subscales appear to be on the same scale? **[YOUR ANSWER]**
:::

---

# Part 4: Descriptive Statistics and Visualisation

::: {.callout-note}
## Slides Part 4
Complete this section during **Part 4: Descriptives & Visualisation** practice time.
:::

## Overall Descriptives

```{r}
#| label: overall-descriptives

# Calculate overall mean, SD, and n for each scale
cat("=== Consistency of Interest ===\n")
cat("M =", round(mean(grit_data$consistency_interest, na.rm = TRUE), 2), "\n")
cat("SD =", round(sd(grit_data$consistency_interest, na.rm = TRUE), 2), "\n")
cat("n =", length(grit_data$consistency_interest), "\n\n")

cat("=== Perseverance of Effort ===\n")
cat("M =", round(mean(grit_data$perseverance_effort, na.rm = TRUE), 2), "\n")
cat("SD =", round(sd(grit_data$perseverance_effort, na.rm = TRUE), 2), "\n")
cat("n =", length(grit_data$perseverance_effort), "\n\n")

cat("=== Total Grit ===\n")
cat("M =", round(mean(grit_data$total_grit, na.rm = TRUE), 2), "\n")
cat("SD =", round(sd(grit_data$total_grit, na.rm = TRUE), 2), "\n")
cat("n =", length(grit_data$total_grit), "\n")
```

## Descriptives by Study Year

```{r}
#| label: grouped-descriptives
#| code-annotations: above

# Calculate descriptives by Study_Year
#| 1. Pipe data forward
grit_summary <- grit_data |>
  #| 2. Group by Study_Year
  group_by(Study_Year) |>
  #| 3. Calculate mean, SD, and n for each subscale and total
  summarise(
    n = n(),
    consistency_M = mean(consistency_interest, na.rm = TRUE),
    consistency_SD = sd(consistency_interest, na.rm = TRUE),
    perseverance_M = mean(perseverance_effort, na.rm = TRUE),
    perseverance_SD = sd(perseverance_effort, na.rm = TRUE),
    total_M = mean(total_grit, na.rm = TRUE),
    total_SD = sd(total_grit, na.rm = TRUE),
    #| 4. Remove grouping structure
    .groups = "drop"
  )

# Display the summary table
grit_summary
```

```{r}
#| label: formatted-table

# Create a nicely formatted table
knitr::kable(grit_summary,
             digits = 2,
             caption = "Descriptive Statistics by Study Year",
             col.names = c("Study Year", "n", "Consistency M", "Consistency SD", 
                          "Perseverance M", "Perseverance SD", "Total M", "Total SD"))
```

::: {.callout-note}
## Exercise 4.1

Use the summary table to answer:

1. Which study year has higher Consistency of Interest? **[YOUR ANSWER]**
2. Which study year has higher Perseverance of Effort? **[YOUR ANSWER]**
3. Are the sample sizes equal? **[YOUR ANSWER]**
:::

## Boxplots

```{r}
#| label: boxplot-consistency
#| fig-width: 8
#| fig-height: 5
#| code-annotations: above

# Create a boxplot for Consistency of Interest
#| 1. Set up plot with aesthetics (x = Study_Year, y = consistency_interest)
ggplot(grit_data, aes(x = factor(Study_Year), y = consistency_interest, fill = factor(Study_Year))) +
  #| 2. Add boxplot geometry
  geom_boxplot(alpha = 0.7) +
  #| 3. Customize labels and colors
  scale_x_discrete(labels = c("1" = "First Year", "2" = "Final Year")) +
  scale_fill_manual(values = c("#9B1B30", "#2E7D32"), guide = "none") +
  #| 4. Apply minimal theme
  theme_minimal() +
  #| 5. Add informative labels
  labs(
    title = "Consistency of Interest by Study Year",
    x = "Study Year",
    y = "Consistency of Interest Score"
  )
```

```{r}
#| label: boxplot-perseverance
#| fig-width: 8
#| fig-height: 5

# Create a boxplot for Perseverance of Effort
ggplot(grit_data, aes(x = factor(Study_Year), y = perseverance_effort, fill = factor(Study_Year))) +
  geom_boxplot(alpha = 0.7) +
  scale_x_discrete(labels = c("1" = "First Year", "2" = "Final Year")) +
  scale_fill_manual(values = c("#9B1B30", "#2E7D32"), guide = "none") +
  theme_minimal() +
  labs(
    title = "Perseverance of Effort by Study Year",
    x = "Study Year",
    y = "Perseverance of Effort Score"
  )
```

```{r}
#| label: boxplot-total
#| fig-width: 8
#| fig-height: 5

# Create a boxplot for Total Grit
ggplot(grit_data, aes(x = factor(Study_Year), y = total_grit, fill = factor(Study_Year))) +
  geom_boxplot(alpha = 0.7) +
  scale_x_discrete(labels = c("1" = "First Year", "2" = "Final Year")) +
  scale_fill_manual(values = c("#9B1B30", "#2E7D32"), guide = "none") +
  theme_minimal() +
  labs(
    title = "Total Grit by Study Year",
    x = "Study Year",
    y = "Total Grit Score"
  )
```

::: {.callout-note}
## Exercise 4.2

Describe what you see in the boxplots:

1. Do the groups appear similar or different for Consistency of Interest? **[YOUR ANSWER]**
2. Do the groups appear similar or different for Perseverance of Effort? **[YOUR ANSWER]**
3. Are there any outliers visible? **[YOUR ANSWER]**
4. Which subscale shows the largest difference between groups? **[YOUR ANSWER]**
:::

---

# Part 5: APA Write-Up and Development

::: {.callout-note}
## Slides Part 5
Complete this section during **Part 5: APA Write-Up** practice time.
:::

## Write Your Results Paragraph

Using the summary statistics from Section 4, write an APA-style paragraph describing the descriptive statistics.

::: {.callout-tip}
## Template

Participants reported [description] levels of grit overall. First-year students (M = ___, SD = ___, n = ___) showed [description] compared to final-year students (M = ___, SD = ___, n = ___) on Consistency of Interest. For Perseverance of Effort, first-year students (M = ___, SD = ___) [description] compared to final-year students (M = ___, SD = ___). Overall grit scores were [description] for first-year students (M = ___, SD = ___) compared to final-year students (M = ___, SD = ___). These descriptive patterns suggest [your interpretation].
:::

**YOUR WRITE-UP HERE:**

[Write your APA-style paragraph here, filling in the statistics from your analysis above]

---

## Proactive Coding Challenge

Now try creating your own analysis. You might:

- Calculate descriptives by Gender
- Create a combined plot showing both subscales
- Calculate the difference between subscales for each participant
- Explore the relationship between age and grit scores

```{r}
#| label: proactive-analysis
#| fig-width: 8
#| fig-height: 5

# YOUR CODE HERE
# Create your own analysis or visualization
# Be creative! What else might be interesting to explore?


```

**Describe what you did and what you found:**

[YOUR DESCRIPTION OF YOUR PROACTIVE ANALYSIS]

# Reflection

::: {.callout-note}
## Exercise 5.1

Answer these reflection questions:

1. **What was the most challenging part of this lab?**

[YOUR ANSWER]

2. **What would you do differently next time?**

[YOUR ANSWER]

3. **What questions do you still have?**

[YOUR ANSWER]
:::

# Session Info

```{r}
#| label: session-info

# Record your R environment for reproducibility
sessionInfo()
```

---

::: {.callout-important}
## Submission Checklist

Before submitting, ensure you have:

- [ ] Replaced "YOUR NAME HERE" with your name
- [ ] Completed all exercises (1.1, 2.1, 3.1, 3.2, 4.1, 4.2, 5.1)
- [ ] Written the APA results paragraph
- [ ] Created a proactive analysis or visualization
- [ ] Answered the reflection questions
- [ ] Rendered to HTML successfully
:::
